Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-06-05 R/SI_Appendix_Datasets.Rmd
revision2: Add Dataset 7 and 8 to SI Appendix material

# Dataset 8

## QTL
```{r}
qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
                              lodcolumn == "dyln_fl50" ~ "Daylength",
                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("50% flowering", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                          "daylength change (s)",
                                          "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
m <- m_gulf_fl_hyp
qtl_df <- qtlplot
subpop = "Gulf"
phe = "flowering"

qtl_mash_overlap_function <- function(m, subpop, phe){

  mash_df <- switchgrassGWAS:::get_marker_df(m) %>%
    mutate(log10BF = switchgrassGWAS:::get_log10bf(m)) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                    sep = "_", extra = "merge", convert = TRUE) %>%
    dplyr::arrange(.data$Chr, .data$Pos) %>%
    select(-value)

  for(i in 1:nrow(qtl_df)){  # from 1 to 16
    ## In block & > 2
    qtl_sub <- mash_df %>%
      filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                               qtl_df$POS_hi[i])) %>% 
      filter(log10BF > 2)

    if(i == 1){
    outputdf <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
                       QTL_Pos_hi = qtl_df$POS_hi[i], 
                       QTL_phenotype = qtl_df$lodcolumn[i], 
                       QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
                       mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
                       mash_phenotype = phe)
    } else {
      outputdf1 <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
                          QTL_Pos_hi = qtl_df$POS_hi[i], 
                          QTL_phenotype = qtl_df$lodcolumn[i], 
                          QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
                          mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
                          mash_phenotype = phe)
      outputdf <- outputdf %>%
        full_join(outputdf1)
    }
  }
  return(outputdf)
}

overlap1 <- qtl_mash_overlap_function(m_gulf_fl_hyp, subpop = "Gulf", 
                                      phe = "flowering") 
overlap2 <- qtl_mash_overlap_function(m_midw_fl_hyp, subpop = "Midwest", 
                                      phe = "flowering") 
overlap3 <- qtl_mash_overlap_function(m_gam_fl_hyp, subpop = "Both", 
                                      phe = "flowering")

dataset_8 <- overlap1 %>%
  full_join(overlap2) %>%
  full_join(overlap3) %>%
  arrange(Chr, mash_Pos)

write_csv(dataset_8, file = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_8_QTL_mash_overlap.csv")
```

Is there an enrichment of SNPs in the top 1% (5%? 10%?) of the distribution for each mash result in the QTL regions?
1% is probably best out of these three arbitrary cutoffs.
```{r}
m <- m_gulf_fl_hyp
qtl_df <- qtlplot

get_QTL_enrichment <- function(m, qtl_df, quantile = 0.95){
  pw_df_all <- get_post_weights(m)

  BFthresh <- quantile(pw_df_all$log10BayesFactor, quantile)
  pw_df <- pw_df_all %>%
  filter(log10BayesFactor > BFthresh) %>%
  select(Chr:Pos, log10BayesFactor, everything())

  for(i in 1:nrow(qtl_df)){  # from 1 to 16
  ## In block & > 2
  QTLandBF <- pw_df %>%
    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                             qtl_df$POS_hi[i])) %>% 
    tally()
  ## BF>2
  inBF <- pw_df %>%
    tally()
  ## BF not > 2 (all minus BF > 2, all is below)
  allwrtBF <- pw_df_all %>%
    tally()
  ## All in block
  inQTL <- pw_df_all %>%
    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                             qtl_df$POS_hi[i])) %>%
    tally()
  notinBF <- (allwrtBF$n[1]-inBF$n[1])
  pval <- phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], 
                 inQTL$n[1], lower.tail = FALSE)

  if(i == 1){
  outputdf <- tibble(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i], inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
  } else {
    outputdf <- outputdf %>%
      add_row(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i],inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
    }
  }
  return(outputdf)
}

get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      dplyr::mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    dplyr::left_join(log10bf_df, by = "Marker") %>%
    dplyr::arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

qtlplot <- qtlplot %>%
  arrange(desc(lod)) %>%
  mutate(pos = as.integer(pos)) %>%
  group_by(CHR, pos) %>%
  slice(1) %>%
  arrange(CHR, pos)
enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Gulf", phenotype = "flowering")
enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Midwest", phenotype = "flowering")
enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Both", phenotype = "flowering")

enrich_df <- enrich_gf %>%
  full_join(enrich_mf) %>%
  full_join(enrich_gamf) %>%
  arrange(Chr, Pos_lo)
enrich_df <- enrich_df %>%
  left_join(qtlplot, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
  select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
  unique() %>%
  arrange(Chr, Pos_lo, lodcolumn)

enrich_df %>%
  filter(pvalue < 0.05)

t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
qtlplot %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.

```

16 QTL. In all cases there was at least one SNP in the QTL interval that was in the 1% tail of SNPs in one of the three flowering mash runs.
We also looked for enrichment of SNPs in the 1% tail of significance in the mash distribution within each QTL interval. (in random genomic intervals of the same size do we see a similar level of enrichment?)
Half of these QTL had a significant enrichment of significant results for mash compared to expectations given the number of SNPs in the QTL interval.

2K: 2 QTL (lod 14 & 17), 1 enriched (in Midwest) (lod 17: CGDD_12C)
2N: 4 QTL (3 non-overlapping) (lod 19, 39, 17, 15), 2 enriched (lod 39 dylnfl50 in Midwest, in Both & 15 CGDD_12C in all three)
4K: 2 QTL, 2 enriched (lod 40 dyln in Midwest & 33 dyln in Gulf)
5N: 4 QTL (3 non-overlapping) (lod 55, 25, 17, 33), 3 enriched (2 non-overlapping) (lod 25 dyln in Midwest, 17 CGDD in Gulf, in Midwest, 33 dyln in Gulf, in Both)
7K: 2 QTL, 0 en (lod 14, 13)
8N: 1 QTL, 0 en (lod 15)
9N: 1 QTL, 0 en (lod 18)

>>>>

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("AnnotationDbi")
BiocManager::install("GenomicFeatures")
BiocManager::install("VariantAnnotation")
```

Find: `qtlplot` in Analysis_v0.6_QTL_interval_plotting. Really, find code to remake this with new QTL from revision2.

Find this file on NAS: "~/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite"

Couldn't find it on the NAS, scp'd it from the Juenger Lab Server:
`scp -i ~/Documents/server_keys/jls_alice alice@popgen.biosci.utexas.edu:/home/alice/Github/pvdiv-genome/Pvirgatum_516_v5.1.gene.txdb.sqlite ~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/`

It's too big to commit; had to add it to data-raw/ignore files ignored by .gitignore.

```{r}
qtldf <- read_csv(here("phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv"))

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  dplyr::select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "Flowering Date",
                         lodcolumn == "dyln_fl50" ~ "Daylength",
                         lodcolumn == "dyln_change_sec" ~ 
                           "daylength change (s)",
                         lodcolumn == "dyln_change_sec_2d_prior" ~ 
                           "daylength change, 2 days prior (s)",
                         lodcolumn == "GR50" ~ "Greenup Date",
                         lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("Flowering Date", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                 "daylength change (s)",
                                 "daylength change, 2 days prior (s)",
                                 "Rainfall, 1 day",
                                 "Greenup Date",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
library(AnnotationDbi)
library(GenomicFeatures)
library(switchgrassGWAS)
qtl_anno_df <- qtlplot %>%
  dplyr::rename(start = POS_lo, end = POS_hi)
not_all_na <- function(x) any(!is.na(x))
txdb <- loadDb("~/Documents/Github/pvdiv-phenology-gxe/data-raw/ignore/Pvirgatum_516_v5.1.gene.txdb.sqlite")
# source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")

ftanno <- read_csv(here("data/Pvirgatum_v5.1.annotation_FTGeneHomologs.csv"))
ftkey <- read_delim(here("data", "OsGene_keyword_table.txt"),
                    delim = "\t")
ftanno %>%
  filter(!is.na(OsGene)) %>%
  left_join(ftkey, by = c("OsGene" = "Symbol")) %>%
  filter(Keyword %in% c("flowering time", "heading date")) %>% unique()

osgene_keyword_df <- ftkey %>%
  dplyr::select(-RAPdb, -MSU) %>%
  group_by(Symbol, Keyword) %>%
  dplyr::slice(1) %>%
  pivot_wider(names_from = Keyword, values_from = Title)

osgene_key <- osgene_keyword_df %>%
  dplyr::select(Symbol, `flowering time`, `heading date`, anther, flower,
                `floral meristem`, `floral meristem determinacy`, flowering,
                `flower development`, `carpel differentiation`, stamen, 
                `stamen number`, vegetative, `plant growth`, `chloroplast development`, chloroplast) %>%
  pivot_longer(cols = -Symbol, names_to = "Keyword", values_to = "Title") %>%
  filter(!is.na(Title)) %>%
  pivot_wider(names_from = Keyword, values_from = Title)

os_annos <- ftanno %>%
  filter(OsGene %in% osgene_key$Symbol | !is.na(AtGene)) %>%
  mutate(Pos_Mb = round(start/10000)/100,
         #Pos_Mb_e = floor(start/10000)/100,
         POS_bin = round(Pos_Mb*100/2)) %>%# ,
         #POS_bin_hi = round(Pos_Mb_e*100/2))
  dplyr::select(OsGene, AtGene, locusName, everything()) %>%
  dplyr::rename(`Gene ID` = locusName)

qtl_annos <- switchgrassGWAS::pvdiv_table_topsnps(df = qtl_anno_df, type = "table", txdb = txdb) %>%
    inner_join(os_annos, by = c("Gene ID")) %>%
    left_join(osgene_key, by = c("OsGene" = "Symbol")) %>%
    dplyr::select(where(not_all_na)) %>%
    group_by(CHR, pos, `Gene ID`) %>%
    dplyr::slice(1)

names(qtl_annos)
all_na_gr_fl <- rowSums(is.na(qtl_annos[,59:70]))
add_column(qtl_annos, GR_FL_Osgene_NA = all_na_gr_fl, .before = TRUE) %>%
  group_by(CHR, pos, GR_FL_Osgene_NA < 12) %>%
  tally()
flowering_qtl_annos <- add_column(qtl_annos, GR_FL_Osgene_NA = all_na_gr_fl, .before = TRUE) %>%
  filter(GR_FL_Osgene_NA < 12) %>%
  dplyr::select(CHR, pos, phe, lod, Pos_Mb_lo, Pos_Mb_hi, ci_lo, ci_hi, `Gene ID`,  gene_start:gene_width, `Arabidopsis thaliana homolog`:`Rice gene annotation`, OsGene, anther:`floral meristem determinacy`) |> 
  mutate(
    Phenology_type = case_when(phe %in% 
                                 c("Flowering Date", 
                                   "daylength change, 2 days prior (s)") ~
                                 "Flowering Date",
                               phe %in% c("Greenup Date", 
                                          "daylength change (s)") ~ 
                                 "Greenup Date",
                               TRUE ~ NA_character_
                               ),
    .before = 3)

flowering_qtl_annos |> 
  filter(Phenology_type == "Greenup Date")

flowering_qtl_annos |> 
  arrange(Phenology_type) |> 
  write_csv(here("manuscript/revision2/SI_Appendix/Dataset_7_QTL_intervals_with_functionally_validated_homologs.csv"))
```

# Dataset 8
All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor > 2, or in the 1% tail of significance, whichever was stricter.

## Objects generated earlier
Get `qtlplot` object by running code starting at about line 150 (for Dataset 7).
Get `m_gulf_fl_hyp` and other objects by running code starting at about line 35 (for Datasets 1-6).

## Thresholds
Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF\>2, whichever is stricter

1% tail is stricter in all cases except Midwest flowering, where BF\>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).

What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF \> 2?

```{r}
#quantile(pw_gf$log10BayesFactor, 0.99)  # Gulf: Use 2.56 BF
#quantile(pw_mf$log10BayesFactor, 0.99)  # Midwest: Use 2 BF
#quantile(pw_gamf$log10BayesFactor, 0.99)  # GAM: Use 2.65 BF
#quantile(pw_mf$log10BayesFactor, 0.995119)

#quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989 BF
#quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831 BF
#quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865 BF
```

Now there are significant greenup date QTL, separate QTL & mash into greenup and flowering associations, and look for overlaps just within each phenology type.

```{r}
m <- m_gulf_fl_hyp
qtl_df <- qtlplot |>   # split by Phenology type. Define based on Figure 4.
  mutate(
    Phenology_type = case_when(phe %in% 
                                 c("Flowering Date", 
                                   "daylength change, 2 days prior (s)") ~
                                 "Flowering Date",
                               phe %in% c("Greenup Date", 
                                          "daylength change (s)") ~ 
                                 "Greenup Date",
                               TRUE ~ NA_character_
                               ),
    .before = 3)
qtl_fl <- qtl_df |> 
  filter(Phenology_type == "Flowering Date")
qtl_gr <- qtl_df |> 
  filter(Phenology_type == "Greenup Date")

subpop = "Gulf"
phe = "Flowering Date"

qtl_mash_overlap_function <- function(qtl, m, subpop, phe, onepcttail = 2){

  mash_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      dplyr::mutate(value = row_number()) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value) |> 
  ## TODO: FIX list/matrix column here
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                    sep = "_", extra = "merge", convert = TRUE) %>%
    dplyr::arrange(.data$Chr, .data$Pos)

  for(i in 1:nrow(qtl)){  # from 1 to 16
    ## In block & > 2
    qtl_sub <- mash_df %>%
      filter(Chr %in% qtl$CHR[i] & between(Pos, qtl$POS_lo[i],
                                               qtl$POS_hi[i])) %>% 
      filter(log10BayesFactor > 2)

    if(i == 1){
    outputdf <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                       QTL_Pos_hi = qtl$POS_hi[i], 
                       QTL_phenotype = qtl$lodcolumn[i], 
                       QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                       mash_log10BayesFactor = qtl_sub$log10BayesFactor, mash_subpop = subpop, 
                       mash_phenotype = phe)
    } else {
      outputdf1 <- tibble(Chr = qtl$CHR[i], QTL_Pos_lo = qtl$POS_lo[i], 
                          QTL_Pos_hi = qtl$POS_hi[i], 
                          QTL_phenotype = qtl$lodcolumn[i], 
                          QTL_LOD = qtl$lod[i], mash_Pos = qtl_sub$Pos, 
                          mash_log10BayesFactor = qtl_sub$log10BayesFactor,
                          mash_subpop = subpop, 
                          mash_phenotype = phe)
      outputdf <- outputdf %>%
        full_join(outputdf1)
    }
  }
  outputdf <- outputdf |> 
    mutate(
      Marker_in_mash_1pct_tail = 
        case_when(mash_log10BayesFactor >= onepcttail ~ TRUE,
                  mash_log10BayesFactor < onepcttail ~ FALSE,
                  TRUE ~ NA)
    )
  return(outputdf)
}

overlap1 <- qtl_mash_overlap_function(qtl_fl, m_gulf_fl_hyp, 
                                      subpop = "Gulf", 
                                      phe = "Flowering Date",
                                      onepcttail = 2.56) 
overlap2 <- qtl_mash_overlap_function(qtl_fl, m_midw_fl_hyp, 
                                      subpop = "Midwest", 
                                      phe = "Flowering Date",
                                      onepcttail = 2) 
overlap3 <- qtl_mash_overlap_function(qtl_fl, m_gam_fl_hyp, 
                                      subpop = "Both", 
                                      phe = "Flowering Date",
                                      onepcttail = 2.65) 
overlap4 <- qtl_mash_overlap_function(qtl_gr, m_gulf_gr_hyp, 
                                      subpop = "Gulf", 
                                      phe = "Greenup Date",
                                      onepcttail = 2.0989) 
overlap5 <- qtl_mash_overlap_function(qtl_gr, m_midw_gr_hyp, 
                                      subpop = "Midwest", 
                                      phe = "Greenup Date",
                                      onepcttail = 4.6831) 
overlap6 <- qtl_mash_overlap_function(qtl_gr, m_gam_gr_hyp, 
                                      subpop = "Both", 
                                      phe = "Greenup Date",
                                      onepcttail = 3.0865)

overlap4 %>%
  full_join(overlap5) %>%
  full_join(overlap6) |> 
  arrange(Chr, mash_Pos)
dataset_8 <- overlap1 %>%
  full_join(overlap2) %>%
  full_join(overlap3) %>%
  full_join(overlap4) %>%
  full_join(overlap5) %>%
  full_join(overlap6) |> 
  arrange(mash_phenotype, Chr, mash_Pos)

write_csv(dataset_8, file = here("manuscript", "revision2", "SI_Appendix",
                                 "Dataset_8_QTL_mash_overlap.csv"))
```

### 2024-06-05 manuscript/revision2/SI_Appendix/SI_Appendix.qmd
revision2: Add Dataset 7 and 8 to SI Appendix material

@@

Canonical covariance matrices can be created by *mash* and fall into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices. The identity matrix has values of zero for covariance between different gardens and correlations of one within gardens, and represents the situation in which the effects in different conditions are independent. The equal effects matrix has values of one for all matrix elements, and represents identical effects across all conditions. Each singleton matrix has a value of one for a single garden and zeros for all other elements, representing a garden-specific effect. Three simple heterozygousity matrices are generated, which have covariances of 0.25, 0.5, or 0.75 for all off-diagonal matrix elements, and correlations of one for diagonal entries, representing moderately positively correlated effects across all sites.

@@

We defined distinct sets of weather-based covariance matrices for the two phenological dates and for the three genetic subgroups. We used the same five weather variables to transform both green-up and flowering date, but defined the weather-based covariance matrices relative to each of the two phenological dates. Weather-based covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.

@@

@@

>>>>

A trait measured in two environments can be considered two traits that are genetically correlated [@falconer1952]. By extension, a phenological trait measured at eight common gardens can be modeled as eight genetically correlated traits represented by a covariance matrix, where diagonal entries represent narrow-sense heritability at a common garden, or the additive genetic variance present in the population at that garden, and non-diagonal entries represent the genetic correlation between the trait in a pair of environments.

We were interested in the types of genetic correlation that could be expressed by different alleles across the genome and reasoned that different loci could have different patterns of genetic correlation. For example, one locus's effects could covary with a weather-based cues, if it had an allele that was responsive to this cue. Alternatively, a locus might have an effect at only one garden, for example if a herbivore or pathogen was only present at one garden.

@@

Canonical covariance matrices were created by *mash* and fell into four groups: an identity matrix, an equal effects matrix, singleton matrices, and simple heterozygousity matrices. The identity matrix has values of zero for covariance between different gardens and correlations of one within gardens, and represents the situation in which the effects in different conditions are independent. The equal effects matrix has values of one for all matrix elements, and represents identical effects across all conditions. Each singleton matrix has a value of one for a single garden and zeros for all other elements, representing a garden-specific effect. Three simple heterozygousity matrices are generated, which have covariances of 0.25, 0.5, or 0.75 for all off-diagonal matrix elements, and correlations of one for diagonal entries, representing moderately positively correlated effects across all sites.

@@

We defined distinct sets of GxWeather covariance matrices for the two phenological dates and for the three genetic subgroups. We used the same five weather variables to transform both green-up and flowering date, but defined the weather-based covariance matrices relative to each of the two phenological dates. GxWeather covariance matrices were also defined separately for individual genotypes from the Gulf, Midwest, and Both subpopulations.

@@

We generated 48 - 60 GxWeather covariance matrices for each of our five weather variables over different time frames. These GxWeather covariance matrices were in many cases highly correlated.

selection of the GxWeather

> describing that the weather-based cues are themselves correlated, and picking and choosing among weather cues is a model building problem – to address the issue, we used a heuristic greedy algorithm to pick the weather-based covariance structures that improved the likelihood of the mash model

> A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices. This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution. If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data. The mash algorithm doesn’t give the full posterior distribution on the loadings so you can’t check for posterior correlations there. It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up.

@@

\dataset{Dataset_7_QTL_intervals_with_functionally_validated_homologs.csv}{Genes in the quantitative trait loci (QTL) regions shown in Figure 4 that have functionally validated homologs in rice for vegetative or reproductive development. The first 9 columns of this table provide QTL information, the remainder provide annotation information for the switchgrass genes in the QTL interval with homologs in A. thaliana or rice that have functional validation in flowering or vegetative-growth related traits. Columns 20-30 provide the titles of papers providing functional validation of the rice homolog of the switchgrass gene.}

\dataset{Dataset_8_blah.csv}{Overlap between quantitative trait loci (QTL) for the start of vegetative or reproductive growth with significant mash effect estimates (with a log10(Bayes Factor) > 2) found using diversity panels of the Midwest, Gulf, and Both genetic subpopulations. The first five columns of this table provide QTL information, and columns 6-10 provide mash Marker information for markers within the QTL regions.}
